---
title: "EDLD 610 Winter 2020 Final Project"
author: "Jim Wright"
date: "2/3/2020"
output: html_document
theme: united 
highlight: zenburn 
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      dev.args = list(bg="transparent"))

library(tidyverse)
library(rio)
library(here)
library(colorblindr)
library(gghighlight)
library(forcats)
library(ggrepel)
library(gt)
library(knitr)
library(kableExtra)


all_ages <- import(here("data", "all-ages.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 


grad_students <- import(here("data", "grad-students.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

majors_lists <- import(here("data", "majors-list.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

recent_grads <- import(here("data", "recent-grads.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

women_stem <- import(here("data", "women-stem.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

```

# Plot 1
## Data is ranked in order of median income with the highest incomes at the top of the data set 
### The purpose of these plots will be to display the distribution of median salaries by college major, by specifically focusing on both the top 20 median salaries and bottom 20 median salaries
```{r plot 1, include=TRUE}

head(recent_grads)
head(women_stem)

recent_grads %>% 
  select(major, median, p25th, p75th)


ggplot(recent_grads, aes(major_category, median)) +
  geom_col() +
  coord_flip()
#need to average the median salary per specific major category levels 

ggplot(recent_grads, aes(fct_reorder(major_category, median), median)) +
  geom_col() +
  geom_col(fill = "cornflowerblue",
  data = filter(recent_grads, major_category == "Education")) +  
  coord_flip()

ggplot(recent_grads, aes(fct_reorder(major, median), median)) +
  geom_col(aes(fill = major_category),
  alpha = 0.7) +
  scale_fill_viridis_d() +
  coord_flip()


median_top_20 <- recent_grads %>% 
  arrange(desc(median)) %>% 
  slice(1:20)

ggplot(median_top_20, aes(fct_reorder(major, median), median)) +
  geom_col(aes(fill = major_category),
  alpha = 0.7) +
  geom_text(aes(major, median, label = scales::dollar(median)),
            nudge_y = 0.02,
            size = 3) +
  scale_fill_brewer(palette = "Dark2") +
  coord_flip() +
  scale_y_continuous("Median Salary", 
                     labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "College Major",
       y = "Median Salary",
       fill = "Major Category",
       title = "Top 20 Median Salaries by College Major") 

median_bottom_20 <- recent_grads %>% 
  arrange(median) %>% 
  slice(1:20)

ggplot(median_bottom_20, aes(fct_reorder(major, median), median)) +
  geom_col(aes(fill = major_category),
  alpha = 0.7) +
  geom_text(aes(major, median, label = scales::dollar(median)),
            nudge_y = 0.02,
            size = 3) +
  scale_fill_brewer(palette = "Dark2") +
  coord_flip() +
  scale_y_continuous("Median Salary", labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "College Major",
       y = "Median Salary",
       fill = "Major Category",
       title = "Bottom 20 Median Salaries by College Major") 


ggplot(recent_grads, aes(median)) +
  geom_histogram(fill = "#56B4E9",
                 color = "black", 
                 alpha = 0.9,
                 bins = 25) +
  theme_minimal(base_size = 15) +
  scale_x_continuous("Median Salary of College Majors", labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  labs(x = "Median Salary of College Majors",
       y = "Total",
       title = "Distribution of Median Salaries per College Major")

p <- ggplot(recent_grads, aes(median, major_category)) +
  ggridges::geom_density_ridges(color = "white",
                                fill = "#A9E5C5") +
  theme_minimal(base_size = 10) +
  scale_x_continuous("Median Salary of College Majors", labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  labs(x = "Median Salary Distribution", 
       y = "Major Category",
       title = "Distribution of Median Salaries by Major Category")

ggplot(recent_grads, aes(median, major_category)) +
  ggridges::geom_density_ridges(color = "white",
                                fill = "#A9E5C5") +
  theme_minimal(base_size = 10) +
  scale_x_continuous("Median Salary of College Majors", labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  labs(x = "Median Salary Distribution", 
       y = "Major Category",
       title = "Distribution of Median Salaries by Major Category")
#Having trouble adding geom_text to ggridges 

```

# Plot 2
## The purpose of Plot 2 will be to communicate visualizations of the recent-grads data to display trends in unemployment rate per college major and major category 
```{r plot 2, include=TRUE}

percentages <- recent_grads %>% 
  mutate(percent_men = men/total, 
         percent_women = women/total, 
         total_jobs = college_jobs + non_college_jobs + low_wage_jobs, 
         percent_college_jobs = college_jobs/total_jobs,
         percent_non_college_jobs = non_college_jobs/total_jobs, 
         percent_low_wage_jobs = low_wage_jobs/total_jobs,
         unemployment_rate_percentage = unemployment_rate)

percents_tidy <- percentages %>% 
  pivot_longer(
    cols = c(22:23),
    names_to = "sex",
    values_to = "sex_percentages"
  ) 

percentages %>% 
  summarize(mean(median))

percentages %>% 
  summarize(mean(unemployment_rate_percentage))

plot_2_a <- percents_tidy %>% 
  filter(median >= 40000,
         unemployment_rate_percentage <= 6.50)

plot_2_b <- percents_tidy %>% 
  filter(median <= 40000,
         unemployment_rate_percentage >= 6.50)

ggplot(plot_2_a, aes(unemployment_rate_percentage, median, color = major_category)) +
  geom_point(size = 2,
             alpha = 0.5) +
  scale_color_viridis_d(name = "Major Category") +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5),
        legend.position = "bottom") +
  scale_x_log10("Unemployment Rate Percentage", labels = scales::percent) +
  scale_y_log10("Median Salary", labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  labs(x = "Unemployment Rate Percentage",
       y = "Median Salary",
       title = "Comparison of Employment Rate to Median Salary",
       subtitle = "College Majors with Above Average Median Salary and Below Average Unemployment Rate") 


ggplot(plot_2_b, aes(unemployment_rate_percentage, median, color = major_category)) +
  geom_point() +
  theme_minimal() +
  scale_color_viridis_d(name = "Major Category") +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5),
        legend.position = "bottom") +
  scale_x_log10("Unemployment Rate Percentage", labels = scales::percent) +
  scale_y_log10("Median Salary", labels = scales::dollar) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  labs(x = "Unemployment Rate Percentage",
       y = "Median Salary",
       title = "Comparison of Employment Rate to Median Salary",
       subtitle = "College Majors with Below Average Median Salary and Above Average Unemployment Rate") 



ggplot(percents_tidy, aes(fct_reorder(major_category, unemployment_rate), unemployment_rate)) +
  geom_col(fill = "cornflowerblue",
           alpha = 0.7) +
  geom_text(aes(major_category, unemployment_rate, label = paste0(round(unemployment_rate), "%",
            size = 2))) +
  coord_flip() +
  scale_y_continuous("Unemployment Rate", labels = scales::percent) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "Major Category",
       title = "Unemployment Rates per Major Category") 
#need to keep playing with this one

plot_2_c <- percents_tidy %>% 
  select(major, major_category, unemployment_rate_percentage, percent_college_jobs) %>% 
  arrange(desc(unemployment_rate_percentage)) %>% 
  slice(1:20)

ggplot(plot_2_c, aes(unemployment_rate_percentage, percent_college_jobs, color = major_category))+
  geom_point(size = 2) +
  scale_color_viridis_d(name = "Major Category") +
  geom_text_repel(aes(label = major),
                  size = 3) +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5),
        legend.position = "bottom") +
  scale_x_log10("Unemployment Rate", labels = scales::percent) +
  scale_y_log10("Employment in College Major", labels = scales::percent) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  labs(x = "Unemployment Rate",
       y = "Employment in College Major",
       title = "Comparison of Employment Rate to Employment in College Major",
       subtitle = "College Majors with the Highest Unemployment Rate") 

```


# Plot 3
## The purpose of plot 3 will be to use the recent_grads and women_stem data sets to compare trends in the percentage of women working in these fields 
```{r plot 3, include=TRUE}
head(recent_grads)
head(women_stem)


stem_percentages <- women_stem %>% 
  mutate(percent_men = men/total,
         percent_women = women/total)

stem_tidy <- stem_percentages %>% 
  pivot_longer(
    cols = c(10:11),
    names_to = "sex",
    values_to = "sex_percentages"
  )

engineering <- stem_percentages %>% 
  filter(major_category == "Engineering")

engineering_2 <- stem_tidy %>% 
  filter(major_category == "Engineering")

ggplot(engineering, aes(percent_men, percent_women)) +
  geom_point(aes(color = major))

ggplot(engineering, aes(fct_reorder(major, percent_women), percent_women)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous("Percentage of Women Employed", labels = scales::percent) 
  
ggplot(engineering, aes(fct_reorder(major, percent_women), percent_women)) +
  geom_col(fill = "#0000FF",
          alpha = 0.7) +
  geom_text(aes(major, percent_women, label = scales::percent(percent_women, digits = 2)),
            nudge_y = -0.05,
            size = 3,
            color = "white") +
  coord_flip() +
  scale_y_continuous("Percentage of Women Employed", labels = scales::percent) 


ggplot(engineering_2, aes(fct_reorder(major, sex_percentages), sex_percentages)) +
  geom_col(fill = "cornflowerblue",
           alpha = 0.7,
           width = 1) +
  coord_flip() +
  facet_wrap(~sex, ncol = 1) +
  geom_text(aes(label = round(sex_percentages), 2)) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "Sex Percentage",
       y = "Major",
       title = "Sex Percentage of Enginnering Majors")

  
top_10 <- stem_percentages %>% 
  group_by(major_category) %>% 
  top_n(2, percent_women)

bottom_10 <- stem_percentages %>% 
  group_by(major_category) %>% 
  top_n(-2, percent_women)

ggplot(stem_percentages, aes(percent_women, percent_men)) +
  geom_point(color = "gray80") +
  geom_point(color = "red", data = bind_rows(top_10, bottom_10)) +
  geom_text_repel(aes(label = major),
                  data = bind_rows(top_10, bottom_10),
                  size = 2) +
  facet_wrap(~major_category) +
  theme_minimal() +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5)) +
  scale_x_log10("Percent Women Employed", labels = scales::percent) +
  scale_y_log10("Percent Men Employed", labels = scales::percent) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_line(color = "gray80")) +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5)) +
  labs(title = "Gender Employment Percentages in STEM Related Fields")

table <- bind_rows(top_10, bottom_10) %>% 
  select(major, median, percent_women, percent_men) 


table %>% 
  gt() %>% 
  cols_label(major = "College Major",
             median = "Median Salary",
             percent_women = "Percent Women Employed",
             percent_men = "Percent Men Employed") %>% 
  cols_align(align = "left", columns = vars(major)) %>% 
  cols_align(align = "center", columns = vars(percent_men, percent_women)) %>% 
  tab_header(title = "STEM Major Statistics") %>% 
  fmt_percent(vars(percent_women, percent_men), decimals = 0) %>% 
  fmt_currency(vars(median), currency = "USD") %>% 
  cols_align(align = "left", columns = vars(major)) %>% 
  tab_source_note(source_note = md("Data from [fivethirtyeight](https://https://github.com/fivethirtyeight/data/tree/master/college-majors)"))




#info_currencies()
```

