---
title: "EDLD 610 Winter 2020 Final Project"
author: "Jim Wright"
date: "2/3/2020"
output: html_document
theme: united 
highlight: zenburn 
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      dev.args = list(bg="transparent"))

library(tidyverse)
library(rio)
library(here)
library(colorblindr)
library(gghighlight)
library(forcats)
library(ggrepel)


all_ages <- import(here("data", "all-ages.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 


grad_students <- import(here("data", "grad-students.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

majors_lists <- import(here("data", "majors-list.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

recent_grads <- import(here("data", "recent-grads.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

women_stem <- import(here("data", "women-stem.csv"),
               setclass = "tbl_df") %>% 
               janitor::clean_names() 

```

# Plot 1
## Data is ranked in order of median income with the highest incomes at the top of the data set 
### The purpose of this plot will be to display the distribution of median salaries by college major, by specifically focusing on both the top 20 median salaries and bottom 20 median salaries
```{r plot 1, include=TRUE}

head(recent_grads)
head(women_stem)

recent_grads %>% 
  select(major, median, p25th, p75th)


ggplot(recent_grads, aes(major_category, median)) +
  geom_col() +
  coord_flip()
#need to figure out how to average the median salary per specific major category levels 

ggplot(recent_grads, aes(fct_reorder(major_category, median), median)) +
  geom_col() +
  geom_col(fill = "cornflowerblue",
  data = filter(recent_grads, major_category == "Education")) +  
  coord_flip()

ggplot(recent_grads, aes(fct_reorder(major, median), median)) +
  geom_col(aes(fill = major_category),
  alpha = 0.7) +
  scale_fill_viridis_d() +
  coord_flip()


median_top_20 <- recent_grads %>% 
  arrange(desc(median)) %>% 
  slice(1:20)

ggplot(median_top_20, aes(fct_reorder(major, median), median)) +
  geom_col(aes(fill = major_category),
  alpha = 0.7) +
  scale_fill_OkabeIto() +
  coord_flip() +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "College Major",
       y = "Median Salary",
       fill = "Major Category",
       title = "Top 20 Median Salaries by College Major") 

median_bottom_20 <- recent_grads %>% 
  arrange(median) %>% 
  slice(1:20)

ggplot(median_bottom_20, aes(fct_reorder(major, median), median)) +
  geom_col(aes(fill = major_category),
  alpha = 0.7) +
  scale_fill_OkabeIto() +
  coord_flip() +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "College Major",
       y = "Median Salary",
       fill = "Major Category",
       title = "Bottom 20 Median Salaries by College Major") 

recent_grads %>% 
  count(major)

median_middle_40 <- recent_grads %>% 
  arrange(median) %>% 
  slice(80:120)

ggplot(median_middle_40, aes(fct_reorder(major, median), median)) +
  geom_col(aes(fill = major_category),
  alpha = 0.7) +
  scale_fill_viridis_d(option = "magma") +
  coord_flip() +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "College Major",
       y = "Median Salary",
       fill = "Major Category",
       title = "Bottom 20 Median Salaries by College Major") 

ggplot(recent_grads, aes(median)) +
  geom_histogram(fill = "#56B4E9",
                 color = "black", 
                 alpha = 0.9,
                 bins = 25) +
  theme_minimal(base_size = 15) +
  labs(x = "Median Salary of College Majors",
       y = "Total",
       title = "Distribution of Median Salaries per College Major")

ggplot(recent_grads, aes(median, major_category)) +
  ggridges::geom_density_ridges(color = "white",
                                fill = "#A9E5C5") +
  theme_minimal(base_size = 10) +
  labs(x = "Median Salary Distribution", 
       y = "Major Category",
       title = "Distribution of Median Salaries by Major Category")

```

# Plot 2
## The purpose of Plot 2 will be to communicate visualizations of the recent-grads data to display trends in unemployment rate per college major and major category 
```{r plot 2, include=TRUE}

percentages <- recent_grads %>% 
  mutate(percent_men = men/total*100, 
         percent_women = women/total*100, 
         total_jobs = college_jobs + non_college_jobs + low_wage_jobs, 
         percent_college_jobs = college_jobs/total_jobs*100,
         percent_non_college_jobs = non_college_jobs/total_jobs*100, 
         percent_low_wage_jobs = low_wage_jobs/total_jobs*100,
         unemployment_rate_percentage = unemployment_rate*100)

percents_tidy <- percentages %>% 
  pivot_longer(
    cols = c(22:23),
    names_to = "sex",
    values_to = "sex_percentages"
  ) 

percentages %>% 
  summarize(mean(median))

percentages %>% 
  summarize(mean(unemployment_rate_percentage))

plot_2_a <- percents_tidy %>% 
  filter(median >= 40000,
         unemployment_rate_percentage <= 6.50)

plot_2_b <- percents_tidy %>% 
  filter(median <= 40000,
         unemployment_rate_percentage >= 6.50)

ggplot(plot_2_a) +
  geom_point(aes(unemployment_rate_percentage, median, color = major_category)) +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "Unemployment Rate Percentage",
       y = "Median Salary",
       title = "Comparison of Employment Rate to Median Salary",
       subtitle = "College Majors with Above Average Median Salary and Below Average Unemployment Rate") 

ggplot(plot_2_a, aes(unemployment_rate_percentage, median)) +
  geom_point(aes(color = major_category)) +
   geom_label_repel(aes(label = major_category),
                   box.padding = 0.35,
                   point.padding = 0.5,
                   segment.color = "grey50") +
  theme_minimal() +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "Unemployment Rate Percentage",
       y = "Median Salary",
       title = "Comparison of Employment Rate to Median Salary",
       subtitle = "College Majors with Above Average Median Salary and Below Average Unemployment Rate") 

ggplot(plot_2_b, aes(unemployment_rate_percentage, median)) +
  geom_point(aes(color = major_category), position = "jitter") +
  geom_label_repel(aes(label = major_category),
                   box.padding = 0.35,
                   point.padding = 0.5,
                   segment.color = "grey50") +
  theme_minimal() +
  theme(plot.title = element_text(color = "black", 
                                  size = 12, 
                                  face = "bold", 
                                  hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "Unemployment Rate Percentage",
       y = "Median Salary",
       title = "Comparison of Employment Rate to Median Salary",
       subtitle = "College Majors with Below Average Median Salary and Above Average Unemployment Rate") 
#like the idea but needs a lot of cleaning up and trimming of included data 


ggplot(percents_tidy, aes(fct_reorder(major_category, unemployment_rate), unemployment_rate)) +
  geom_col(fill = "cornflowerblue",
           alpha = 0.7) +
  coord_flip() +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5),
        legend.position = "bottom") +
  labs(x = "Unemployment Rate",
       y = "Major Category",
       title = "Unemployment Rates per Major Category") 



```


# Plot 3
## The purpose of plot 3 will be to use the recent_grads and women_stem data sets to compare trends in the percentage of women working in these fields 
```{r plot 3, include=TRUE}
head(recent_grads)
head(women_stem)


stem_percentages <- women_stem %>% 
  mutate(percent_men = men/total*100,
         percent_women = women/total*100)

stem_tidy <- stem_percentages %>% 
  pivot_longer(
    cols = c(10:11),
    names_to = "sex",
    values_to = "sex_percentages"
  )

engineering <- stem_percentages %>% 
  filter(major_category == "Engineering")

engineering_2 <- stem_tidy %>% 
  filter(major_category == "Engineering")

ggplot(engineering, aes(percent_men, percent_women)) +
  geom_point(aes(color = major))

ggplot(engineering, aes(fct_reorder(major, percent_women), percent_women)) +
  geom_col() +
  coord_flip() 
  

ggplot(engineering_2, aes(fct_reorder(major, sex_percentages), sex_percentages)) +
  geom_col(fill = "cornflowerblue",
           alpha = 0.7,
           width = 1) +
  coord_flip() +
  facet_wrap(~sex, ncol = 1) +
  geom_text(aes(label = sex_percentages)) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "Sex Percentage",
       y = "Major",
       title = "Sex Percentage of Enginnering Majors")

  
 

```

